\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-06-05 europa-basics.tex
updates

\begin{figure}[b]
\centering
  \includegraphics[scale=0.45]{figs/shopping_europa.pdf}
  \caption{\small A potential plan for a robot shopping for groceries.}
\label{fig:shop:shopping}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[scale=0.3]{figs/Example-NQueens1.jpg}
\caption{\small An $N$-Queens solution generated by \eue.}
\label{fig:nqueens-2}
\end{figure}

@@

\begin{figure}[t]
\centering
\subfloat[\small A \eu solution to an \texttt{RCPSP} \cite{Bruckera99}
  problem.]{\label{fig:rcpsp-1}\includegraphics[scale=0.25]{figs/Example-UBO0.jpg}}\qquad
\subfloat[\small A \eu solution to a shopping agent problem domain where
  the agent needs to buy Bananas, Milk and a Drill.]{\label{fig:shopping-1}\includegraphics[scale=0.3]{figs/Example-Shopping0.jpg}}
\caption{\small \eu based solutions to scheduling and planning problems.}
\end{figure}

@@

\begin{figure}[!t]
\centering
\includegraphics[scale=0.3]{figs/constraint-net.pdf}
\caption{\small Variables and Constraints represented as a Constraint
  Network \kcomment{before arc consistency is enforced}.}
\label{fig:constraintnet}
\end{figure}

>>>>

\begin{figure}[b]
\centering
  \includegraphics[scale=3]{figs/shopping_europa.jpeg}
  \caption{\small A potential plan for a robot shopping for groceries.}
\label{fig:shop:shopping}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[scale=2.0]{figs/Example-NQueens1.jpeg}
\caption{\small An $N$-Queens solution generated by \eue.}
\label{fig:nqueens-2}
\end{figure}

@@

\begin{figure}[t]
\centering
\subfloat[\small A \eu solution to an \texttt{RCPSP} \cite{Bruckera99}
  problem.]{\label{fig:rcpsp-1}\includegraphics[scale=1.8]{figs/Example-UBO0.jpeg}}\qquad
\subfloat[\small A \eu solution to a shopping agent problem domain where
  the agent needs to buy Bananas, Milk and a Drill.]{\label{fig:shopping-1}\includegraphics[scale=2.2]{figs/Example-Shopping0.jpeg}}
\caption{\small \eu based solutions to scheduling and planning problems.}
\end{figure}

@@

\begin{figure}[!t]
\centering
\includegraphics[scale=1.9]{figs/constraint-net.jpeg}
\caption{\small Variables and Constraints represented as a Constraint
  Network \kcomment{before arc consistency is enforced}.}
\label{fig:constraintnet}
\end{figure}

### 2012-06-05 europa.tex
updates

\begin{figure}[b]
\centering
\includegraphics[scale=0.5]{figs/EUROPA-Architecture.pdf}
\caption{\small The \eu software architecture.}
\label{fig:europa-architecture}
\end{figure}

@@

\kcomment{Before going into the details of each module, a simple
  example illustrates how they interact at an abstract level:}

\begin{verbatim}

class LightBulb extends Timeline
{
    predicate On  {}
    predicate Off {}
}

class LightSwitch extends Timeline
{
    LightBulb myBulb_;

    LightSwitch(LightBulb b)
    {
        myBulb_ = b;
    }

    action turnOn  { duration=1; }
    action turnOff { duration=1; }
}

LightSwitch::turnOn
{
    met_by(condition object.myBulb_.Off);  // Bulb must be Off to be turned On
    meets(effect object.myBulb_.On);   // Must be turned on through the switch
}

LightSwitch::turnOff
{
    met_by(condition object.myBulb_.On);   // Bulb must be On to be turned Off
    meets(effect object.myBulb_.Off);      // Must be turned off through the switch
}

// Problem instance : turning the light off
LightBulb bulb1 = new LightBulb();
LightSwitch switch1 = new LightSwitch(bulb1);

@@

// We want the bulb to be off by time 10
goal(bulb1.Off goal1);
lt(0,goal1.start);
lt(goal1.start,10);

\end{verbatim}

\kcomment{The example above is a \nd (covered in more detail in
  Section \ref{sec:europa:modeling}) model of a light bulb and a
  switch that controls it along with a problem instance that requires
  the light bulb to be \texttt{off} by time$=10$, given an initial
  state with the bulb \texttt{on} at time$=0$. The \nd module allows
  \eu to ingest this model and the problem description. Once the \nd
  statements are parsed, different object types, object instances and
  constraints are instantiated inside the \texttt{PDB} and
  \texttt{CRE} modules where they are available for the user and the
  solver modules to query and manipulate programmatically. This
  instantiation of the initial state constitutes the initial plan that
  \eu holds for this particular problem instance.

@@

To articulate the most important elements of \nde, we show how the
Shopping Agent problem could be modeled:

\begin{verbatim}

// Locations (Home, SuperMarket, etc.)
class Location {
  string name;
  Location(string _name){
    name = _name;
  }
}

// Products (Milk, Banana, etc.) and
class Product {
  string name;
  Product(string _name) {
    name = _name;
  }
}

// ProductLocations (Banana can be found at SuperMarket, for example)
class ProductLocation {
  Location location;
  Product product;

  ProductLocation(Location _location, Product _product){
    location = _location;
    product = _product;
  }
}

// Use built-in Timeline functionality to enforce that an agent:
// a) Can't be at more than one place at a time.
// b) Can't Go more than one place at a time.
// c) Can't Go somewhere and be At somewhere at the same time.
class AgentLocation extends Timeline{
  predicate At {
    Location loc;
  }

  predicate Going {
    Location from;
    Location to;
  }
}

// In addition to having a location timeline, the agent can buy and own things.  
// Note that the actions and the location predicates can be concurrent 
// so they can't be on the same Timeline
class Agent {
  AgentLocation location;

  Agent() {
    location = new AgentLocation();
  }

  action Buy {
    Product product;
  }

  action Go {
    Location from;
    Location to;
  }

  predicate Own {
  	Product product;
  }
}
\end{verbatim}

The model above the main elements of a problem domain can be mapped to
\eus representation in the \kcomment{following manner}:

@@

\begin{verbatim}

// Define the rules for our actions:
Agent::Go {
  met_by(condition object.location.At origin);
  eq(from, origin.loc);

@@

  meets(effect Own purchase);
  eq(purchase.product,product);
}
\end{verbatim}

Domain rules can be defined in \nd in the context of a predicate or an
action. In the example above, the main elements of a rule definition
\kcomment{are as follows}:

@@

With the specification for the Shopping Agent domain \kcomment{above},
a particular problem instance could be stated as:

\begin{verbatim}

@@

\end{verbatim}

This problem instance consists of \textbf{Variables} and
\textbf{Objects}, \textbf{Facts}, \textbf{Goals} and
\textbf{Constraints}. In our example we define temporal variables that
are used to specify the planning horizon (\texttt{horizonStart} and
\texttt{horizonEnd}) and objects are used to specify locations,
products, product-location associations and the Shopping Agent that we
want to plan for. Facts, represented as tokens in the \texttt{PDB},
are used to specify the current state of the world \kcomment{and} do
not need any support. In this \kcomment{example} we use facts to
specify that the Shipping Agent is initially at home. Goals are also
represented as tokens in the \texttt{PDB}. Different from facts, goals
trigger inference in \eu which determine whether flaws (unresolved
subgoals, overloaded resources, etc.) need to be resolved; if so,
search modules can be invoked to resolve those flaws as part of the
deliberation process. Finally, constraints are restrictions on and
between facts and goals. In the example above, we use temporal
constraints to specify when facts hold and when goals need to be
accomplished. \texttt{before} and \texttt{after} are \texttt{Allen
  Algebra} \cite{allen84} operators, which \eu supports to specify
temporal relations between tokens (specifically between their start
and end timepoints). The algebra is particularly effective in dealing
with interval arithmetic; its elements are summarized in
Fig. \ref{fig:allen-algebra}

\begin{figure}[!t]
\centering
\includegraphics[scale=0.4]{figs/Allen-algebra.pdf}
\caption{\small Temporal relations defined within the planner are
  based on \texttt{Allen Algebra} \cite{allen84} relations shown
  above.}
\label{fig:allen-algebra}
\end{figure}

@@

\begin{figure}[!htb]
  \centering
  \subfloat[\small A Simple Temporal Network (STN). \kcomment{Nodes correspond
  to variables and their domains and arc's the \emph{constraint} or
  distance between them.}]{\label{fig:stn1}\includegraphics[width=0.4\columnwidth]{figs/stn-graph.pdf}}
  \subfloat[\small The resulting distance graph from \ref{fig:stn1}.]{\label{fig:stn2}\includegraphics[width=0.4\columnwidth]{figs/distance-graph.pdf}}
  \caption{\small An illustration of a Simple Temporal Network
    (\texttt{STN}) and its corresponding distance graph.}
  \label{fig:stn}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[scale=0.2]{figs/europa-resource-envelopes.png}
\caption{\small An example of resource envelopes \cite{Muscettola04,Muscettola06} computed by \eu.}
\label{fig:resenvelopes}
\end{figure}

>>>>

\begin{figure}[b]
\centering
\includegraphics[scale=2.5]{figs/EUROPA-Architecture.jpeg}
\caption{\small The \eu software architecture.}
\label{fig:europa-architecture}
\end{figure}

@@

Before going into the details of each module, a simple example
illustrates how they interact at an abstract level.

\begin{figure}[h!]
\begin{verbatim}

class LightBulb extends Timeline
{
    predicate On  {}
    predicate Off {}
}
class LightSwitch extends Timeline
{
    LightBulb myBulb_;

    LightSwitch(LightBulb b)
    {
        myBulb_ = b;
    }
    action turnOn  { duration=1; }
    action turnOff { duration=1; }
}
LightSwitch::turnOn
{
    met_by(condition object.myBulb_.Off);  // Bulb must be Off to be turned On
    meets(effect object.myBulb_.On);   // Must be turned on through the switch
}
LightSwitch::turnOff
{
    met_by(condition object.myBulb_.On);   // Bulb must be On to be turned Off
    meets(effect object.myBulb_.Off);      // Must be turned off through the switch
}
// Problem instance : turning the light off
LightBulb bulb1 = new LightBulb();
LightSwitch switch1 = new LightSwitch(bulb1);

@@

// We want the bulb to be off by time 10
goal(bulb1.Off goal1);
lt(0,goal1.start);
lt(goal1.start,10);
\end{verbatim}
\caption{\small \nd modeling example for a light bulb and switch control.}
\label{fig:nddllight}
\end{figure}

\kcomment{Fig. \ref{fig:nddllight} is a \nd (covered in more detail in
  Section \ref{sec:europa:modeling}) model of a light bulb and a
  switch that controls it along with a problem instance that requires
  the light bulb to be \texttt{off} by time$=10$, given an initial
  state with the bulb \texttt{on} at time$=0$. The \nd module allows
  \eu to ingest this model and the problem description. Once the \nd
  statements are parsed, different object types, object instances and
  constraints are instantiated inside the \texttt{PDB} and
  \texttt{CRE} modules where they are available for the user and the
  solver modules to query and manipulate programmatically. This
  instantiation of the initial state constitutes the initial plan that
  \eu holds for this particular problem instance.

@@

\begin{figure}[tbp]
\begin{verbatim}
// Locations (Home, SuperMarket, etc.)
class Location {
  string name;
  Location(string _name){
    name = _name;
  }
}
// Products (Milk, Banana, etc.)
class Product {
  string name;
  Product(string _name) {
   name = _name;
  }
}
// ProductLocations (Banana can be found at SuperMarket, for example)
class ProductLocation {
   Location location;
   Product product;

   ProductLocation(Location _location, Product _product){
     location = _location;
     product = _product;
   }
}
// Use build-in timeline functionality to enforce that an agent cannot
// be at more than one place at one time
// Go to more than one place at one time
// Go somewhere and be At somewhere at the same time 
class AgentLocation extends Timeline{
  predicate At {
    Location loc;
  }
  predicate Going {
    Location from;
    Location to;
   }
}
// In addition to having a location timeline, the agent can buy and
// own things. Note that the actions and the location predicates can
// be concurrent so they cannot be on the same timeline.
class Agent {

  AgentLocation location;
  Agent() {
   location = new AgentLocation();
  }
  action Buy {
   Product product;
  } 
  action Go {
   Location from;
   Location to;
  }
  predicate Own {
   Product product;
  }
}
\end{verbatim}
\caption{\small \nd model of the Shopping Agent example.}
\label{fig:nddlshop}
\end{figure}

To articulate the most important elements of \nde, we show how the
Shopping Agent problem could be modeled in
Fig. \ref{fig:nddlshop}. The main elements of a problem domain can be
mapped to \eus representation in the \kcomment{following manner}:

@@

\begin{figure}[h!]
\begin{verbatim}

Agent::Go {
  met_by(condition object.location.At origin);
  eq(from, origin.loc);

@@

  meets(effect Own purchase);
  eq(purchase.product,product);
}
\end{verbatim}
\caption{\small \nd rules for the Shopping Agent.}
\label{fig:nddlshop-rules}
\end{figure}

Domain rules can be defined in \nd in the context of a predicate or an
action. In Fig. \ref{fig:nddlshop-rules}, the main elements of a rule
definition \kcomment{are as follows}:

@@

With the specification for the Shopping Agent domain \kcomment{above},
a particular problem instance could be stated as in
Fig. \ref{fig:nddlshop-instance}.

\begin{figure}[h!]
\begin{verbatim}

@@

\end{verbatim}
\caption{\small \nd instance and goal specification for the Shopping Agent.}
\label{fig:nddlshop-instance}
\end{figure}

This problem instance consists of \textbf{Variables} and
\textbf{Objects}, \textbf{Facts}, \textbf{Goals} and
\textbf{Constraints}. In our example we define temporal variables that
are used to specify the planning horizon (\texttt{horizonStart} and
\texttt{horizonEnd}) and objects are used to specify locations,
products, product-location associations and the Shopping Agent that we
want to plan for. Facts, represented as tokens in the \texttt{PDB},
are used to specify the current state of the world \kcomment{and} do
not need any support. In this \kcomment{example} we use facts to
specify that the Shipping Agent is initially at home. Goals are also
represented as tokens in the \texttt{PDB}. Different from facts, goals
trigger inference in \eu which determine whether flaws (unresolved
subgoals, overloaded resources, etc.) need to be resolved; if so,
search modules can be invoked to resolve those flaws as part of the
deliberation process. Finally, constraints are restrictions on and
between facts and goals. In the example above, we use temporal
constraints to specify when facts hold and when goals need to be
accomplished. \texttt{before} and \texttt{after} are \texttt{Allen
  Algebra} \cite{allen84} operators, which \eu supports to specify
temporal relations between tokens (specifically between their start
and end timepoints). The algebra is particularly effective in dealing
with interval arithmetic; its elements are summarized in
Fig. \ref{fig:allen-algebra}.

\begin{figure}[!t]
\centering
\includegraphics[scale=2.3]{figs/Allen-algebra.jpeg}
\caption{\small Temporal relations defined within the planner are
  based on \texttt{Allen Algebra} \cite{allen84} relations shown
  above.}
\label{fig:allen-algebra}
\end{figure}

@@

\begin{figure}[!htb]
  \centering
  \subfloat[\small A Simple Temporal Network (\texttt{STN}). \kcomment{Nodes correspond
  to variables and their domains and arc's the \emph{constraint} or
  distance between them.}]{\label{fig:stn1}\includegraphics[width=0.5\columnwidth]{figs/stn-graph.jpeg}}
  \subfloat[\small The resulting distance graph from \ref{fig:stn1}.]{\label{fig:stn2}\includegraphics[width=0.5\columnwidth]{figs/distance-graph.pdf}}
  \caption{\small An illustration of a \texttt{STN} and its
    corresponding distance graph.}
  \label{fig:stn}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[scale=2]{figs/europa-resource-envelopes.jpeg}
\caption{\small An example of resource envelopes \cite{Muscettola04,Muscettola06} computed by \eue.}
\label{fig:resenvelopes}
\end{figure}

### 2012-06-05 future.tex
updates

Our experience of using deliberative techniques in a highly
inter-disciplinary environment for targeted marine science
applications has shown the need and applicability of such novel
algorithms and methods. In addition to augmenting traditional AUV
surveys more advanced Lagrangian observations, mixed-initiative
methods, our colleagues at MBARI have understood the longer term
implications of our work. Our current and consequently future efforts
are therefore grounded in important science problems in an operational
oceanographic setting.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.45]{figs/autonomy-topics.pdf}
  \caption{\small Overlapping topics in our research in deliberation
    and autonomy with sampling the dominant theme.}
  \label{fig:topics}
\end{figure}

The core of our problem domain for \can has been towards Sampling. In
large part given the spatial and temporal scales of the
bio-geochemical processes we are interested in in this project, it is
imperative that a key direction is in formulating a multi-robot
adaptive sampling problem. In this context while superficially \rx can
be considered as a multi-agent system and could therefore imply a
direct extension, sporadic communication with high-latencies in
surface communication, not to mention low throughput acoustic
communications imply that the extension of \rx for distributed
applications is non-trivial. Further, in order to efficiently execute
coordinated observation of ocean processes with multiple autonomous
sensing platforms, significant methodological as well as technical
challenges must be addressed. We are working towards a systematic
approach of using \rx as a backend planner for \can field
experiments. The expectation is that \od users will plan initially
individual vehicle surveys which are variants of standard
straight-line transects. Lessons learned with human planners in the
loop will be used towards planning larger surveys, also with simple
fixed transects but using vehicle capabilities in ontologies similar
to \cite{patron09}. Apportioning the planning task between shore-side
components with substantial contextual information and compute power
with robots with in-situ reasoning capabilities a la \rx is a
challenge and an open research problem. Fig. \ref{fig:topics} shows
what we believe is the scope and influence of the different topic
areas we are focusing on.

Partitioned inference and decision-making within \rx also lends itself
to graceful system degradation, making the overall approach effective
towards increased robustness. There are two outstanding diagnosis and
failure recovery challenges that we need to address; the first has to
do with methodology related to model design of the instrument payload
for a highly modular and configurable vehicle; this is local to a
reactor. Doing so would enable (as a first step) using systematic
Fault Diagnosis, Isolation and Recovery (FDIR) algorithms as
demonstrated in the \texttt{RAX} experiment
\cite{williams96,mus98,williams97} and more recently on an AUV by
Dearden and others \cite{wang09,ernits10,dearden11}. The second
challenge has to \kcomment{do with} implementing a policy within the
\rx framework that explicitly deals with recovery since controller
failure can be distributed across reactors. Since reactors are
hierarchical in their relationship to one another, their well defined
dependencies allow them to be removed from an agent
systematically. Reactors more abstract in scope can be removed before
those which are less abstract (and closer to the hardware). In the
event a Navigation reactor, for example, has a fault, it can be
removed safely with lower level functionalities (such as those in the
executive). When the Navigator is removed, it will recursively remove
more abstract reactors (such as a Mission Manager) that depend on
it. Such a safety feature then allows the system designers to ensure
that early on they can invest more in ensuring that reactors lower in
the hierarchy are robust and potentially formally verified. It also
means that there is a safe and sure way to take control of the vehicle
at any level within the reactors hierarchy; as long as low level
reactors are still active one will always been able to send
goals/commands to them and execute what are traditionally known as
runout sequences (or clean up scripts).

Feature recognition for Sampling is a general research thrust; however
it has specific relevance in the context of \rx and AUV control. As
noted in Section \ref{sec:results} we have already demonstrated
Machine Learning techniques for recognition of select features in the
coastal ocean. A principal challenge that remains is that of obtaining
expert labeled data for any supervised or semi-supervised learning
methods.  While supervised techniques for learning classifiers such as
decision trees (\cite{Quinlan93-dtrees}), instance-based
classification (\cite{Aha-ibl-ml91}), Bayesian classifiers
(\cite{Jensen2001-BNetworks}) and artificial neural networks
(\cite{ANNsurvey-2000}) are popular, given the vast amount of data
available and poor understanding of correlation between physical and
biogeochemical variables in the coastal ocean, our work on INLs
\cite{mcgann08d,ryan10} and reliance on a oceanographic expert has
shown that scaling to different problems remains problematic. This is
a prime motivation to move towards more semi-supervised methods such
as \cite{kumar11}. With the advent of the \od however,
another direction our research is taking us is in \texttt{Recommender
  Systems} \cite{Adomavicius05} as noted in Section
\ref{sec:results}. We are working to design and deploy software
infrastructure that will learn from user tagged data. To do so, we
will use tagging techniques similar to what users of commercial image
sharing sites like Flickr and Picassa do, to learn the context of the
data (including remote sensing). In addition, the \od will integrate
statistical Machine Learning techniques with incoming asynchronous
data stream obtained by sensors and platforms. Identification models
will be based on existing training sets for INLs and plumes; models
for other features of scientific interests driven by \can requirements
will be built in addition. Existing AUV archives for instance provide
sufficient data for a learning base for a range of features of
interest to \can; in addition we will build software that will learn
to incorporate incoming data streams to provide an existence proof of
a feature of interest.

@@

More abstractly at the architectural level, the ability to dynamically
manipulate the reactor graph would be of importance while creating new
reactors or modifying timelines. Doing so would provide more robust
failure recovery mechanisms especially after a reactor has failed to
synchronize. For example, an alternate reactor could take ownership of
timelines left invalid by the destruction of another failed reactor
ensuring a transparent implementation of component redundancy. While
the current design broadly provides for such a capability, the \rx API
has yet to implement it while ensuring other design elements within
\rx and \eu are not compromised.

>>>>

Our experience of using deliberative techniques in a highly
inter-disciplinary environment for targeted marine science
applications has shown the need and applicability of such novel
algorithms and methods. In addition to augmenting traditional AUV
surveys, more advanced Lagrangian observations and mixed-initiative
methods, we believe our colleagues at MBARI have understood the longer
term implications of our work. Our current and consequently future
efforts are therefore grounded in important science problems in an
operational oceanographic setting.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.45]{figs/autonomy-topics.jpeg}
  \caption{\small Overlapping topics in our research in deliberation
    and autonomy with sampling the dominant theme.}
  \label{fig:topics}
\end{figure}

The core of our problem domain for \can has been towards Sampling. In
large part given the spatial and temporal scales of the
bio-geochemical processes we are interested in, in this project, it is
imperative that a key direction is in formulating a multi-robot
adaptive sampling problem. In this context while \rx can be considered
as a 'multi-agent' system and \kcomment{could therefore imply its use
  in a multi-vehicle setting as} a direct extension, sporadic
communication with high-latencies in surface communication, not to
mention low throughput acoustic communications imply that the
extension of \rx for distributed applications is non-trivial. Further,
in order to efficiently execute coordinated observations of ocean
processes with multiple autonomous sensing platforms, significant
methodological as well as technical challenges must be
addressed.

\kcomment{In this context} we are working towards a systematic
approach of using \rx as a backend planner \kcomment{within the \od}
for \can field experiments. The expectation is that \od users will
\kcomment{initially be able to} plan individual vehicle surveys which
\kcomment{would be} variants of straight-line transects. Lessons
learned with human planners in the loop will be used towards planning
larger surveys, also with simple fixed transects but using vehicle
capabilities in ontologies similar to \cite{patron09}. \kcomment{Doing
  so will likely enable the right vehicle to target an ocean feature
  with appropriate sensors and sampling strategy}. Apportioning the
planning task between shore-side components with substantial
contextual information and compute power, with robots having in-situ
reasoning capabilities similar to \rxe, is \kcomment{challenging in of
  itself} and an open research problem. \kcomment{In pursuing this
  line of research, we expect to touch on a range of associated
  problems for decision-theoretic control of vehicles in the coastal
  ocean}. Fig. \ref{fig:topics} shows what we believe is the scope and
influence of the different topic areas we are \kcomment{currently}
focusing on.

Partitioned inference and decision-making within \rx also lends itself
to graceful system degradation, making the overall approach effective
towards increased \kcomment{mission} robustness \kcomment{in the
  context of fault diagnosis, isolation and recovery (FDIR)}. There
are two outstanding diagnosis and failure recovery challenges that we
need to address; the first has to do with methodology related to model
design of the instrument payload for a highly modular and configurable
vehicle; this is local to a reactor. Doing so would enable (as a first
step) using systematic FDIR algorithms as demonstrated in the
\texttt{RAX} experiment \cite{williams96,mus98,williams97} and more
recently on AUVs \cite{wang09,ernits10,dearden11}. The second
challenge has to \kcomment{do with} implementing a policy within the
\rx framework that explicitly deals with recovery since controller
failure can be distributed across reactors. Since reactors are
hierarchical in their relationship to one another, their well defined
dependencies allow them to be removed from an agent systematically as
noted in Section \ref{sec:rx-reactor-failure}. Reactors more abstract
in scope can be removed before those which are less abstract (and
closer to the hardware). In the event a \texttt{Navigator} reactor,
for example, has a fault, it can be removed safely with lower level
functionalities (such as those in the executive). When the
\texttt{Navigator} is removed, it will recursively remove more
abstract reactors (such as a \texttt{Mission Manager}) that depend on
it. Such a safety feature then allows the system designers to ensure
that early on they can invest more in ensuring that reactors lower in
the hierarchy are robust and potentially formally verified. It also
means that there is a safe and sure way to take control of the vehicle
at any level within the reactors hierarchy; as long as low level
reactors are still active one will always been able to send
goals/commands to them and execute what are traditionally known as
runout sequences.

Feature recognition for Sampling is a general research thrust; however
it has specific relevance in the context of \rx and AUV control. As
noted in Section \ref{sec:results} we have already demonstrated
Machine Learning techniques for recognition of select features in the
coastal ocean. A principal challenge that remains is that of obtaining
expert labeled data for any supervised or semi-supervised learning
methods.  While supervised techniques for learning classifiers such as
decision trees (\cite{Quinlan93-dtrees}), instance-based
classification (\cite{Aha-ibl-ml91}), Bayesian classifiers
(\cite{Jensen2001-BNetworks}) and artificial neural networks
(\cite{ANNsurvey-2000}) are popular, given the vast amount of data
available and poor understanding of correlation between physical and
biogeochemical variables in the coastal ocean, our work on INLs
\cite{mcgann08d,ryan10} and reliance on a oceanographic expert has
shown that scaling to different problems remains problematic. This is
a prime motivation to move towards more semi-supervised methods such
as \cite{kumar11}.

With the advent of the \od however, another direction our research is
taking us is in \texttt{Recommender Systems} \cite{Adomavicius05} as
noted in Section \ref{sec:results}. We are working to design and
deploy software infrastructure that will learn from user tagged
data. To do so, we will use tagging techniques similar to what users
of commercial image sharing sites like Flickr and Picassa do, to learn
the context of the data (including remote sensing). In addition, the
\od will integrate statistical Machine Learning techniques with
incoming asynchronous data stream obtained by sensors and
platforms. Identification models will be based on existing training
sets for INLs and plumes; models for other features of scientific
interests driven by \can requirements will be built in
addition. Existing AUV archives for instance provide sufficient data
for a learning base for a range of features of interest to scientists;
in addition we will build software that will learn to incorporate
incoming data streams to provide an existence proof of a feature of
interest.

@@

More abstractly at the architectural level, the ability to dynamically
manipulate the reactor graph would be of importance while creating new
reactors or modifying timelines. Doing so would provide more robust
failure recovery mechanisms especially after a reactor has failed to
synchronize. For example, an alternate reactor could take ownership of
timelines left invalid by the destruction of another failed reactor
ensuring a transparent implementation of component redundancy. While
the current design broadly provides for such a capability, the \rx API
has yet to implement it while ensuring other design elements within
are not compromised.
